---
title: "A Focused Start"
author: "Asher Baer"
output: html_document
---
# The End
I'm sick of guides that don't show where we're going; so, let's start at the end.   
Below is the essence of everything I do in R. Don't worry, I'll talk you through it.  

```{r echo = T, results = 'hide', error=FALSE, warning=FALSE,  message=FALSE,  fig.show='hide'}
library(dplyr)
library(ggplot2)

cars_raw <- read.csv("./Data_To_Use/exampleData.csv")

cars_cleaned <- cars_raw %>% 
  dplyr::rename("car_model" = X) %>%
  dplyr::mutate(cylnder_efficiency = mpg/cyl) %>%
  dplyr::select(car_model, cylnder_efficiency, everything()) 

cars_cleaned %>% ggplot(aes(x = disp , y = cylnder_efficiency)) + geom_point()

cars_cleaned %>% lm(cylnder_efficiency ~ disp, data = .) %>% summary()

cars_cleaned %>% write.csv("./Results/cars_cleaned_customWhateverName.csv")
```
This is the universal high-level workflow : Import -> Clean -> Examine -> Save Results.


What's going on here? Let's talk thru it in order of what I did, and what the code sees.

Step-by-step:   
1. Bring in additional functions (through packages)   
2. Read in data stored locally   
2.5 Save that data   
3. "Clean" the data   
3.5 Do something row-wise to the data   
4. Save the cleaned data   
5. Visualize + Do a statistical test   
6. Save the Cleaned Data   

# In Detail, this time

```{r echo = T, results = 'hide', error=FALSE, warning=FALSE,  message=FALSE,  fig.show='hide'}
##NOTE:: "#" indicates a comment, or text that is ignored. 
# 1. "library" brings in pre-installed packages off functions we will use. Here, they are "dplyr" and "ggplot2"
library(dplyr)
library(ggplot2)

#2. "read.csv" is a function that takes in a "path" to some data. 
## PATH? Relative to the file, what other files to click to get to something. 
## "./" means relative. /Data_To_Use looks in Data_To_Use and "exampleData.csv" is selected

#3. "<-" takes the data from the CSV and saves it into a variable called "cars_raw". We could name this whatever
cars_raw <- read.csv("./Data_To_Use/exampleData.csv")

#Save everything that is done to "cars_cleaned"
cars_cleaned <- cars_raw %>% 
  #Rename the column X to "car_model", then
  dplyr::rename("car_model" = X) %>%
  #Create a new column cylinder_efficiency where each value is the value from the mpg columns divided by the cyl columns, then
  dplyr::mutate(cylnder_efficiency = mpg/cyl) %>%
  #take the car_model column, the cylinder_efficiency column, then saves it (go back to top)
  dplyr::select(car_model, cylnder_efficiency, everything()) 

#take cars_cleaned, then
cars_cleaned %>% 
  #create a plot with the x axis as displ and the y as cylinder_efficiency
  ggplot(aes(x = disp , y = cylnder_efficiency)) +
  #view it as a scatter plot
  geom_point()

cars_cleaned %>% 
  #linear regression with Depedent as cylnder_efficiency and independent as disp
  lm(cylnder_efficiency ~ disp, data = .) %>% 
  #then visualize
  summary()

cars_cleaned %>% 
  #save the results
  write.csv("./Results/cars_cleaned_customWhateverName.csv")
```
