---
title: "1 Set Up"
author: "Asher Baer"
output: html_document
---

### Overview

The goal of this document is to get you set-up so you can start using, and learning R.\
It is important to feel you have a comfortable place to experiment. Let's walk thru it.

### Starting at Nothing

At this point you should have nothing at all installed. Great!

We're going to install two things:    
1. R Programming Language. (A place on your computer that speaks R and runs it)   
2. R Studio, an IDE. (A place for us to write R and run it)   

Go here for R : https://cran.r-project.org/    
Go here for R Studio :   https://posit.co/download/rstudio-desktop/    

##### Install Notes    

I'm currently using R 4.3. ish. Any stable version is good.   

### Part 2: Setting Up the IDE

When you open R it should look something like this : 
```{r}

```

Click File -> New File -> R-Markdown File. 
Hit yes on everything, name it whatever. 

Great, now you should see this:
```{r}

```

### Part 2.1:

Making R-Studio your home.   

### Part 2.2: Intro to Workflow and Where to Think

Briefly, I do most of my work in R as a long script. Instead of using callable .R files, I work in 
noteooks (like this one), where I can code inside "chunks"
To start a chunk type ```{r} everything inbetween will be r code. Everything outside will be a markdown file. 

```{r}
#I am a happy little chunk of r code

chunk_code = 1
```

```{r}
#I am in the next chunk, but am in the global environment
print(chunk_code)

```


