---
title: "2 Quick Overview"
author: "Asher Baer"
output: html_document
---

### Overview

The goal of this document is to give a quick-start with more detail for people who do not typically code.

### Running Things

You can run R-code as either a .R or .Rmd file.\
A .R file is an executable, i.e. it will be called and run line-by-line, usually by itself or in a different program.\
A .Rmd file is a notebook, which lets you have "chunks" of code between text. We will work in .Rmds.\
\
An R-Chunk is everything within 3 ` and {r}.
```{r}
#I am an R chunk!
## If you are confused, the "#" means a comment or something that the code ignores. I can say whatever here.
```
Here, you can run your R-Code. You have the option to run an individual chunk, or all chunks sequentially.

### What to Run
A lot of our work will be broken down into variables and functions.\
Variables: Things that hold data. (nouns)\
Functions: Things that do things to data. (verbs)\

#### Variables
Naturally, R is used for numbers. Let's start here.
"Declare" or make a variable by typing it and giving it a value.\
Give it a value with "=" or with left-arrow key and dash (make a little arrow)
```{r}
##Inside the mystical world of an R-Chunk...
#This variable, called "variable_name_whatever" has the value of 3 stored in it.
variable_name_whatever = 3
#This variable called "x" has the value of 4 stored in it.
x <- 4

# "print" is a function that takes the variable, looks inside, and displays the result
print(variable_name_whatever)
```

Variables persist between R Chunks, and can interact.
```{r}
print(x)
```
Variables can be re-assigned
```{r}
variable_name_whatever <- 5
print(variable_name_whatever)
```

#### Functions
Functions are things you type that take input and give output.\
Their inputs are put inside parenthesis, and in the case of multiple inputs, specified directly.

```{r}
#Print takes one input
print(3)
# "+" is a function that takes 2
print(3+3)
```
Functions can take variables as inputs; their outputs can be stored as new variables
```{r}
#Read this as : create a new variable called "new_variable" then take the number in "variable_name_whatever" add 3, save the result in new_variable
new_variable <- variable_name_whatever + 3
print(new_variable)
```
#### A little bit more detail
Variables themselves, as well as functions, fall into categories. These categories are called types and let us do things.
Without bogging down in the detail, I will list out the basic types (called primitive types).

```{r, results = "hold"}
#Integer: Just a whole number. Can be negative.
3
-3
#Double: Numbers with decimal points
3.0
1.2
-1.3
#"Character" : text
"Hello, I am a list of text, code sees me!"
print("Text me")
#"Logical" or Boolean: a T/F value. 
TRUE
T
```
Additionally, there is also one called "factor" which means a list of possible options. We will talk about this later.

##### Non-Primative Types

###### Lists
A list is simply a series of elements.\
```{r, results = "hold"}
#A list is initlaized using "c()" where everything in the quotes is an element.
I_am_a_list <- c(3, "4", x)
print(I_am_a_list)
```
Some more complex types exist, such as tables or data-frames (tables) that let us do more things.\
\
###### Important one: Data-Frame
```{r}
#A dataframe is simply a table that knows it's columns have types.

column_one = c(3,4)
column_two = c("dog", "cat")

#the data.frame function here takes in two lists, and turns them into columns.
dataframe_thingy <- data.frame(column_one, column_two)
print(dataframe_thingy)
```
Note: You may be reading this thinking, "will I be needing to make data, type it into a list, then use that list to make a data-frame? That seems like a lot of work!!\
Answer: NO!! Or not usually. A dataframe is a table, and naturally, we can bring in data to make the dataframe.

### Pause : Some tips for figuring things out.

```{r}
#You can always use the function "typeof()" to check R thinks a variable is.
typeof("hi")
#You can always use ? or ?? on a function to bring up explanations for how it works and what it expects.
?data.frame #This will only come up in R-Studio under "help"
```

### Doing Real Work
Okay, at this point you should know a few things: \
Variables hold data\
Some variables are basically tables\
Functions do things to data, and expect specific types of data to work\
\
Great, but let's do some examples that show how work actually can be.



