---
title: "2a Overview For Coders"
author: "Asher Baer"
output: html_document
---

### Overview
The goal of this document is to give a quick-start with more detail for people experienced in CS.
Meaing C++, C#, or whatever else.\


#### Variable
Declare a variable by typing anything without quotes.Must not start with a number.\
Must either give it a blank object or simply have an object you want to put into it.\
Use either = or less-than with a minus symbol. 

By default, everything will be in the global scope (unless declared in a function).
```{r}
##Examples
babys_first_variable.periods_mean_nothing = 2
print(babys_first_variable.periods_mean_nothing + 2)

x = babys_first_variable.periods_mean_nothing + 2
print(x + babys_first_variable.periods_mean_nothing)

y<-16
y + x

rm(x)
```

### Types? 
R is functional, but objects can have types. There are a few main ones, that will be familiar and obvious.\
1. Integer\
2. Double (like a float)\
3. Character (not just a single character, it's a string in other languages)\
4. Factor (basically an enum)\
5. logical (boolean)\
6. Others not useful for now(Raw/Complex types)\
```{r}
x = as.integer(1)
typeof(x)
x = 1.5
typeof(x)

typeof(TRUE)

```

### Lists
to make a list, put it inside a c()\ 
so, c(1,2,3) is a list of 3 elements.

indexing starts at 1!
```{r}
x = c("hi", "there")
x[1]
x[2]
x[3]
x[4] = 12
print(x)

```

### Dataframes
This will be your bread and butter doing data-science in R.\
A dataframe is effectively a matrix, spreadsheet, or series of columns put together.\
Each column can have it's own type. Let's look at one of the default ones.
```{r}
print(mtcars)
```
We can treat dataframes as objects and save them in variables. In fact, much of the data-cleaning process
will be done by doing something to a data-frame, then saving the output in a variable.

We can index a dataframe 
```{r}
row_index = 3
column_index = 2

print(mtcars[row_index, column_index])
## we can get the column as a list easily
print(mtcars[, column_index])
#or the row as a 1 x ncol data-frame
print(mtcars[row_index, ])
```
### Functions
functions are defined with "function()" then list the input variables. 
Return values with "return()"
```{r}
#
function_whatever_name <- function(input_whatever_var){
  print(input_whatever_var)
  if(is.numeric(input_whatever_var)){print("I'm a numeric too!")
    }
  return(input_whatever_var)
}

function_whatever_name(2)
```
# Now let's get started


### To the point
This has been a useful starting point looking at base R. However, much is done using a great package called the tidyverse.\


#### Install
To install a package in R, go over to the bottom right panel, click "packages", then "install", then type in "dplyr"

To load a package, type in library(package_name).It will be loaded globally, and gie some conflict warnings.
```{r}
library(dplyr)
```
To call a function from a package, you can specific it using package::function
```{r}
#select takes a dataframe, then keeps some of the columns you selected
dplyr::select(mtcars,1)
```

### Pipes
The last fundamental thing to really know is the pipe operator. Written as %>%, it can be read as "and then"\
Basically, it take the output of the last function and uses it as the input for the next. If you want to specify where 
as an argument it should be used, technically it can be done with the period, .\

```{r}
mtcars %>% dplyr::select(1)
mtcars %>% dplyr::select(.,1)
```
### Some dplyr bread and butter functions
We want to manipulate dataframes since it is our way of storing data, interacting with it, and doing things at scale.\
Below are some key functions I use a lot.\
1. dplyr::mutate -- changes the column or creates a new one if that rows is not there\
2. dplyr::case_when -- if else but can be used in mutate\
3. dplyr::filter -- used to keep rows based on a condition (needs a logical)\
4. dplyr::select -- keeps columns, based on names\
5. dplyr::full_join or dplyr::left_join -- joins two dataframes together\
6. dplyr::summarize -- lets you do operations across all rows. any function, even user defined
7. dplyr::group_by -- decides which rows to treat as the same for various operations, like summarize

```{r}
#save the output in this mt_cars_changed
mt_cars_changed <- mtcars %>%
  #create a new column
  dplyr::mutate(new_col = 1) %>%
  #multiple that column by two, and add the value of vs rowwise
  dplyr::mutate(new_col = new_col*2 + vs) %>%
  #select new_col, mpg, mpg cyl, and all columns that are numeric
  dplyr::select(new_col, mpg, cyl, where(is.numeric)) %>%
  #only keep rows where the mpg > 5
  dplyr::filter(mpg > 5) %>%
  #round the mpg to 2 digits
  dplyr::mutate(mpg = round(mpg, 0)) %>%
  #conver the mpg column into a character
  dplyr::mutate(mpg = as.character(mpg))
print(mt_cars_changed)

mt_cars_mean <- mt_cars_changed %>% 
  #group by the mpg variable
  dplyr::group_by(mpg) %>%
  #take the mean of it and the sum
  summarise(mean_hp = mean(hp), number_of_rows = n())

final_df <- mt_cars_changed %>% 
  dplyr::left_join(mt_cars_mean, by  = c("mpg" = "mpg"))%>%
  #select mpg, mean_hp, number_of_rows, then everything else
  dplyr::select(mpg, mean_hp, number_of_rows, everything())
```
#### Visuals
Base R has some good options with the "plot" function.\
ggplot2 has some great stuff that works with dplyr.

```{r}
library(ggplot2)

final_df %>% 
  ggplot(aes(x = mpg, y = cyl, color = mean_hp)) + geom_point()

```

#### Problem Solving in R
It should be noted any function's documentation can be brought up on the bottom right by typing ?function.\
??function will bring up all instances of it in packages installed, but not currently loaded into the environment.
```{r}
?as.numeric
??as.numeric
```